---
title: "Visualization I"
author: "Kristal Quispe"
date: "10/2/2019"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

## Create Weather Data
```{r load_data, cache =TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## create a ggplot

```{r}
ggplot(weather_df, aes(x= tmin, y=tmax)) +
 geom_point()
```

## alternate way of making plot above

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point()
```


saving initial plots... mostly teacher doesnt do this 

```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y= tmax)) +
  geom_point()

scatterplot
```

adding color...

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point(aes(color = name))

weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4)
##alpha code makes the color dot transparent between 0-1
```

why do aesthetic positions matter?

first
```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4)+
  geom_smooth(se = FALSE)
```

vs.
```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```

make the one graph into three graphs for each weather station

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) + #se = false is to get rid of CI
  facet_grid(~name) #says that names wil be in the columns, ~ is a tilda
```

this is fine, but not interesting

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)

#to make the data points proportional to size of precip and to make the smooth line bigger
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha =.35) +
  geom_smooth(size = 2, se = FALSE)
  
weather_df %>% 
  ggplot(aes(x = date, y = prcp, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

## Some extra stuff

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = FALSE)
## you just get smooth line with out data points if you exclude the geopoint. 
```

2d density

```{r}
# install.packages("hexbin")
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d() # or geom_hex()
```

## More kinds of plots!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, color = name)) +
  geom_histogram()

#color is telling you about what is outside of the bars, fill if for inside

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()

#if you dont like stacked histograms you can do the following 

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

#to make three seperate histograms
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```

density plots!!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

univariate plots!!

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

violin plot!!

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots!!! 

good for bimodal and multi group comparisons

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

##saving a plot

```{r}
ggp_rideg_temp =
  weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_rideg_temp)
```

embedding plots in R markdown

```{r, fig.width = 6}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```


```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```
