---
title: "Visualization II"
author: "Kristal Quispe"
date: "10/3/2019"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
#Code above is for general preferences of all plots going forward. 
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Create Weather Data
```{r load_data, cache =TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## Making new plots

start with an old one

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5)

```

So now we are going to add labels:

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  )

```

lets play around with scales: x axis tick marks etc

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C","-5C", "20C" )
  )

```


If you need a log or square root transfromation
```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C","-5C", "20C" )
  ) +
  scale_y_continuous(
    trans ="sqrt"
  )

#or you can do...
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C","-5C", "20C" )
  ) +
  scale_y_sqrt()
```

## Colors next

```{r}
## to change the default colors ggplot sets use a scale color function...
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(50, 250)
    #name is the key legend name, not the variable name
  )
```

viridis package... gives us more color options

```{r}
ggp_base =
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha =.5) +
  labs(
    title = "Temperture plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE ##variable here is a discrete variable, need to specify this. 
  )

ggp_base
```

## themes
```{r}
## theme has to do with general organization functions, here i am moving legend to the bottom
ggp_base +
  theme(legend.position = "bottom")
```

```{r}
#If you want a black and white back drop
ggp_base +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
#If you want a minimal theme back drop, aka get rid of the black out line in grpah
ggp_base +
  theme_minimal ()+
  theme(legend.position = "bottom")

#theme minimal has to happen before them legened position, b/c theme minimal is applying a specfic lengend. Over arching themes go first aka like theme minimal, and any tweeks, like them lengend position go after. 
```

```{r}
#If you want to drop the legend completely
ggp_base +
  theme_minimal ()+
  theme(legend.position = "none")
```

(where do my warnings come from)
```{r}
weather_df %>% 
  filter(is.na(tmin))
```

## More than one data set

```{r}
central_park =
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)

#we have a plot with just one data set and then stacked another data set into the original data set
```

(brief aside about colors)

```{r}
# regular code
waikiki %>% 
  ggplot(aes(x=date, y =tmax, color= name))+
  geom_point()

#code for assigning the data points a sepcific color, specify color in geomtery that takes this value
waikiki %>% 
  ggplot(aes(x=date, y =tmax))+
  geom_point(color = "blue")
```

## Multi-panel plot

```{r}
ggp_scatter =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()

ggp_density =
  weather_df %>% 
  ggplot(aes(x = tmin))+
  geom_density()

ggp_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax, color = name))+
  geom_boxplot()

ggp_scatter + ggp_density

#This will make a scater plot next to density plus, add additional planel with + and put one on bottom using "/"

ggp_scatter + (ggp_density/ggp_box)

#not only way, can use grid arrange but teach doesnt like it

```

## Data Manipulation

```{r}
ggp_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax, color = name))+
  geom_boxplot()
#change order of factors to change order of plt

weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")) %>%
  ggplot(aes(x = name, y = tmax, color = name))+
  geom_boxplot()
  
```

Reorder instead of rlevel

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)) %>%
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()

#Here we are reordering by tmax, lowest to highest t max
```

## Restructure then plot

```{r}
#Plots for density of min temp of all locations and the max temp of all locations. 
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperture"
  ) %>% 
  ggplot(aes(x = temperture, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(~name) +
  theme(legend.position = "bottom")

```

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(visit = recode(visit, "bl" = "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()

#factor is now going in alphabetical order, prior to tidying the data tideying makes this easier. 
```


litter and pups

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 


fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```

